---
title: "Introducing altair"
description: |
  An R interface to the Altair Python Package
author:
  - name: Ian Lyttle
date: 2018-05-20
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
options(altair.embed_options = altair::vega_embed())
```


Introducing **altair**, an R package to work with the Python package [Altair](https://altair-viz.github.io/), which you can use to build and render [Vega-Lite](https://vega.github.io/vega-lite/) chart-specifications: 

https://vegawidget.github.io/altair 

Vega-Lite offers an implementation of an interactive grammar of graphics. The vocabulary and syntax of this grammar is different from that used by [ggplot2](http://ggplot2.tidyverse.org/). However, the fundamental ideas are very much the same: ggplot2 offers *aesthetics*, *geoms*, and *scales*; Vega-Lite offers *marks*, *encondings*, and *scales*. Vega-Lite also offers interactivity "baked into" the rendering.

### Demonstration

Following this [Altair example](https://altair-viz.github.io/user_guide/selections.html), here's a set of scatterplots with linked brushing. You can click-and-drag to select observations in one chart, and they are highlighted in the other:

<details>
<summary>Code</summary>
```{r echo=TRUE}
library("altair")
vega_data <- import_vega_data()
brush <- alt$selection_interval()
chart_mpg <-
  alt$Chart()$
  mark_point()$
  encode(
    x = "Horsepower:Q",
    y = "Miles_per_Gallon:Q",
    color = alt$condition(brush, "Origin:N", alt$value("lightgray"))
  )$
  properties(selection = brush, width = 250, height = 250) 
chart_disp <- 
  chart_mpg$encode(
    x = "Displacement:Q"
  )
chart <- 
  (chart_mpg | chart_disp)$
  properties(
    data = r_to_py(vega_data$cars()),
    title = "Linked-Brush Scatterplot"
  )
```

```{r eval=FALSE}
chart
```
</details>

```{r echo=FALSE}
chart
```

The interactivity is a fundamental part of Vega-Lite; a interactive chart is rendered and controlled entirely in your broswer.


### Resources

You may be initially interested in the [installation instructions](https://vegawidget.github.io/altair/articles/installation.html), as your biggest hassle may be getting your Python environment set up. 

You may also be interested in a [first example](https://vegawidget.github.io/altair/articles/first-example.html), an [interactive example](https://vegawidget.github.io/altair/articles/interactive.html), and the [interactive section of the gallery](https://vegawidget.github.io/altair/articles/example-gallery-08-interactive-charts.html). 

### Acknowledgements

This approach is distinct from efforts to build a native R interface to Vega-Lite, by [Bob Rudis](https://github.com/hrbrmstr/hrbrmstr) and coworkers: the [**vegalite**](https://github.com/hrbrmstr/vegalite) package, which has inspired this effort.

This package rests on foundations provided by [reticulate](https://rstudio.github.io/reticulate/) and [Altair](https://altair-viz.github.io/). [Jake Vanderplas](https://github.com/jakevdp) has been super supportive and patient in answering Altair questions, which is greatly appreciated.  

It is a lot of fun to work with these folks on this project; this package would not exist without their efforts:

 - [Alicia Schep](https://github.com/AliciaSchep) has contributed the chart-concatenation operators, as well as sorted out many of the Python, JavaScript, and package-API issues.

 - [Haley Jeppson](https://github.com/haleyjeppson) has adapted the [Altair Example Gallery](https://altair-viz.github.io/gallery/index.html#) to R; here's the [first page](https://vegawidget.github.io/altair/articles/example-gallery-01-simple.html) of nine.

 - [Heike Hofmann](https://github.com/heike) has been an invaluable advisor, providing incisive feedback, and insight into the fundamentals of interactive graphics.


